Technical Report

### 2018-03-21 fully-abstract-comp/technical-report.tex
updated definition of 'reasonable' component used in full-abstraction premise

% \[
%   \lrv(H_\sigma,W) =
%   \begin{array}[t]{l}
%     \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
% %
%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\
%                          \sealed{\sigma,((\perm,\lin),\baddr,\eaddr,\aaddr + ?)}\endarray\right)} \;\middle| \;
%     \begin{array}{l}
%       (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
%       \npair{([\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})]} \in \lrp(W)
%     \end{array}
%     \right\}\cup\\
% %
%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,(\baddr,\eaddr))},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)} \;\middle| \;
%     \begin{array}{l}
%       \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
%       \forall n' \leq n \ldotp \forall W' \privft W , W_C,W_\lrm,W_\lrr\ldotp \\
%       \quad \npair[n']{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
%       \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\
%                                                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
%       \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
%       \quad \opc = ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C) \wedge \\
%       \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
%       \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
%       \quad \reg_T(r_\stk) = ((\rw,\linear),\stkb,\baddr-1,\_) \wedge \\
%       \quad W'(r_\var{free}) = \stdreg{[\stkb,\baddr-1]}{\spa} \wedge \\
%       \quad W'(r_\var{priv}) = \stareg[\src{\ms_f},\ms_f]{\spao} \wedge \\
%       \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
%       \quad W_\var{new} = (W' \oplus W_\lrr \oplus W_\lrm \oplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
%       \qquad \Rightarrow \\
%       \qquad \quad \npair[n']{\left(\arraycolsep=0pt\array{l}
%                                      \left(\arraycolsep=0pt\array{l}\ms_S,\reg_S \left[
%       \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0,0\endarray \right]\\
%                                      ,\stk,\ms_\stk \uplus \ms_f
%                                      \endarray \right),\\
%                                              \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0,0\endarray\right]\right)
%                                     \endarray\right)} \in \lro
%     \end{array} 
%     \right\}\cup \\
% %
%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
%     \begin{array}{l}
%       \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
%       \forall W' \pubft W, n' \leq n \ldotp \\
%       \quad \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
%       \qquad \npair[n']{(\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T')} \in \lre(W') \wedge \\
%       \qquad \npair[n']{(\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T)} \in \lre(W')
%     \end{array}
%     \right\}\cup\\
% %
%     % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
%     % \right\} \cup \\
% %
%     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
%     \begin{array}{r l l}
%       \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
%       \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
%     \end{array}
%     \right\} \cup \\
% %
%     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
%     \begin{array}{r l l }
%       \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
%       \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
%       \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
%       \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
%     \end{array}
%     \right\}
%   \end{array}
% \]

% \dominique{22-1-2018: the below is an alternative proposal to the above definition. It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{(\src{\vsc_S},\vsc_T)} \in \lrp(W) \wedge \\
      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
      \quad (\isLinear{\vsc_1} \Rightarrow \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
      \quad (\nonLinear{\vsc_1} \Rightarrow \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))

@@

\begin{definition}[Use return seals only for call]
  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either ($\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$) or $\sigma \in \sigcloss$.
\end{definition}

\begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'} \step \src{\Phi''}
  \]
  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
  \[
    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow (\sigrets \cup \sigcloss) \mathrel{\#} [\sigma_b,\sigma_e]
  \]
and
%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
  \[
    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
  \]
  and
    \begin{multline*}
      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in (\sigrets\cup\sigcloss) \Rightarrow 
      \\  \left(  
      \begin{array}{r l}
        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow (\sigrets\cup\sigcloss) \# [\sigma_b,\sigma_e]) \wedge\\
        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
      \end{array}
      \right)
  \end{multline*}
\end{definition}

%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
\begin{definition}[Never store link seals and code capability]
  A configuration $\src{\Phi}$ is said to not store seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
\[
  \Phi'(r_2) = \seal(\sigma_\baddr,\sigma_\eaddr,\_) \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
\]
and
\[
  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
\]
\end{definition}

@@

\dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures). That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.}
\begin{definition}
  A program $\program$ with memory segment $\ms_\program$ and $\dom(\ms_\program) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
  \begin{itemize}
  \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
  \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
  \end{itemize}
\end{definition}

>>>>

\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      \npair{(\src{\vsc_S},\vsc_T)} \in \lrp(W) \wedge \\
      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
      \quad (\isLinear{\vsc_1} \Rightarrow \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
      \quad (\nonLinear{\vsc_1} \Rightarrow \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))

@@

\begin{definition}[Use return seals only for call]
  \label{def:use-return-seals-call}
  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either ($\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$) or $\sigma \in \sigcloss$.
\end{definition}

\begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
  \label{def:handle-trusted-seals-properly}
  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'} \step \src{\Phi''}
  \]
  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
  \[
    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow (\sigrets \cup \sigcloss) \mathrel{\#} [\sigma_b,\sigma_e]
  \]
and
%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
  \[
    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
  \]
  and
    \begin{multline*}
      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in (\sigrets\cup\sigcloss) \Rightarrow 
      \\  \left(  
      \begin{array}{r l}
        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow (\sigrets\cup\sigcloss) \# [\sigma_b,\sigma_e]) \wedge\\
        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
      \end{array}
      \right)
  \end{multline*}
\end{definition}

%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
\begin{definition}[Never store link seals and code capability]
  \label{def:never-store-seal-code-cap}
  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
\[
  \Phi'(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
\]
and
\[
  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
\]
\end{definition}

@@

\dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures). That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.}
\begin{definition}
  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
  \begin{itemize}
  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
  \item $\plug{\context}{\program}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
  \item $\plug{\context}{\program}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
  \end{itemize}
\end{definition}
